---
title: "FinalProject"
author: "Roy Khoo"
date: "2023-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```
# Data Description

The dataset is obtained from kaggle at https://www.kaggle.com/datasets/salvatorerastelli/spotify-and-youtube?resource=download

Some elements that the dataset observed includes the streams, views, album names and other objective statistic of songs released in spotify along with their youtube counterpart. The data was collected in 7th February, 2023 as stated by the providers of the dataset.

# Data Loading

```{r}
df = read_csv("data\\Spotify_Youtube.csv")

```

# Data Cleaning

In order to get working on the dataset, I will likely changed all variables to their correct form, eg: views as numeric, album type as factor, etc. Additionally, I would like to analyze the dataset alongside their youtube counterpart, thus I would remove any dataset that does not have any collection of youtube data. Similarly, I would check for which column has N/A values and drop where it makes sense. This would involve turning all blank and 0 entries into N/A.

```{r}
str(df)
df_clean = df
df_clean[df == "" | df == " "] = NA
sapply(df_clean, function(x) sum(is.na(x)))
```


